---
title: "Querying the Wang Lab Database"
output: html_document
---

```{r setup, include=FALSE}
#Check if the package is installed, if not, install it
if (!requireNamespace("RPostgres", quietly = TRUE)) {
  install.packages("RPostgres")
}
# Check if the package is installed, if not, install it
if (!requireNamespace("DBI", quietly = TRUE)) {
  install.packages("DBI")
}
# Load required packages
library("RPostgres")
library(DBI)
```


### Establishing Connection to the Wang Lab Database
Establishes a connection to the PostgreSQL database using the provided credentials including the database name, host, user, port, and password.
```{r connection, echo=TRUE}
# Enter the parameters from your database
# Do not expose the database credentials

dbname <- 'experiments_data'
host <- 'ep-dry-bonus-a5fb1uxb.us-east-2.aws.neon.tech'
user <- 'readonly_rice_usda_irri'
port <- '5432'

# Connect to postgres database
con <- DBI::dbConnect(RPostgres::Postgres(), dbname = dbname, host = host, port = port, user = user, password = rstudioapi::askForSecret("password"))
```


### Listing Tables in the Database
Lists all the tables available in the connected database using the dbListTables() function.
```{r}
dbListTables(con)
```


### Querying the main table from Database
Defines a SQL query string to select columns and rows from the table named main_table that meet the specified condition. The query is then sent to the database using dbSendQuery(), and the fetched data is stored in a data frame using dbFetch(). Clears the result set using dbClearResult() and disconnects from the database using dbDisconnect() to free up resources. The `data_frame` object remains in memory for analysis even after disconnecting from database.

NOTE: Before executing a new query, RUN the dbClearResult(query) instruction to clean memory from the previous query. Additionally, you should run the dbClearResult(query) before closing the database connection. 

For complex queries, we suggest downloading a subset of data to filter and manipulate the data frame obtained using the functions of R. 
```{r}
# Define the SQL query

# Example query 1: this query returns all records from the table main_table
#query_string <- "SELECT * FROM public.irri_main_table"

# Example query 2: this query returns all records where crop is Wheat
query_string <- "SELECT * FROM public.irri_main_table WHERE treatment = 'Well-Watered'"

# Example query 3: this query returns all records where crop is Wheat and measurement_method is Remote sensing
#query_string <- "SELECT * FROM public.irri_main_table WHERE treatment = 'Drought' AND season = '2023' AND genotype = 'OS 16'"

# Send query to the database
query <- dbSendQuery(con, query_string)

# Fetch the data into a data frame
data_frame <- dbFetch(query)

# Clear previous query
dbClearResult(query)

```


### Querying the management table from Database
Defines a SQL query string to select columns and rows from the table named management_data that meet the specified condition.
```{r}
# Example query 1:  Retrieve all records from the table management_data
query_string <- "SELECT * FROM management_data"

# Send query to the database
query <- dbSendQuery(con, query_string)

# Fetch the data into a data frame
data_frame_mgmt <- dbFetch(query)

# Clear previous query
dbClearResult(query)

```


### Close the database conection
Disconnect from the database to release resources.
```{r}
dbDisconnect(con)
```


### Displaying main table dataframe
Displays the fetched data stored in the `data_frame` object. Data are ready for visualization and analysis!
```{r}
# The results are stored in the data_frame object
head(data_frame)
```


#### Calculate Days After Seeding (DAS)
Extract dates for selected events from the `data_frame_mgmt` table, compute the difference in days relative to dates in the `data_frame` table, and add this difference as a new column.
```{r}
library(dplyr)
library(lubridate)

# Define the name of the day's difference column 
event_name <- 'DAS' 

# Define the event or list of events from the data_frame_mgmt table that are going to be used for the date extraction
selected_events <- c('Seeds sown')

# Convert date columns to Date type
data_frame <- data_frame %>% mutate(date = as.Date(date))

data_frame_mgmt <- data_frame_mgmt %>% mutate(date = as.Date(date))

# Define a function to find the appropriate date from data_frame_mgmt
find_matching_date <- function(row, df_mgmt, selected_events) {
  matching_row <- df_mgmt %>%
    filter(experiment == row$experiment,
           crop == row$crop,
           treatment == row$treatment,
           season == row$season,
           event %in% selected_events)
  
  if (nrow(matching_row) > 0) {
    return(matching_row$date[1])
  } else {
    return(NA)
  }
}

# Apply the function to each row in data_frame to find the matching date
data_frame <- data_frame %>%
  rowwise() %>%
  mutate(matching_date = find_matching_date(pick(everything()), data_frame_mgmt, selected_events)) %>%
  ungroup()

# Calculate the difference in days and add it as a new column EVENT
data_frame <- data_frame %>%
  mutate(!!event_name := as.integer(difftime(date, matching_date, units = "days")))

# Drop the 'matching_date' column as it's no longer needed
data_frame <- data_frame %>%
  select(-matching_date)

# View the resulting data_frame
print(data_frame)

```
