---
title: "Querying the Wang Lab Database"
output: html_document
---

```{r setup, include=FALSE}
#Check if the package is installed, if not, install it
if (!requireNamespace("RPostgres", quietly = TRUE)) {
  install.packages("RPostgres")
}
# Check if the package is installed, if not, install it
if (!requireNamespace("DBI", quietly = TRUE)) {
  install.packages("DBI")
}
# Load required packages
library("RPostgres")
library(DBI)
```


### Establishing Connection to the Wang Lab Database
Establishes a connection to the PostgreSQL database using the provided credentials including the database name, host, user, port, and password.

```{r connection, echo=TRUE}
# Enter the parameters from your database
# Do not expose the database credentials

host <- 'host'
dbname <- 'database_name'
user <- 'user'
port <- 'port_number'

# Connect to postgres database
con <- DBI::dbConnect(RPostgres::Postgres(), dbname = dbname, host = host, port = port, user = user, password = rstudioapi::askForSecret("password"))
```


### Listing Tables in the Database
Lists all the tables available in the connected database using the dbListTables() function.

```{r}
dbListTables(con)
```


### Querying Data from Database
Defines a SQL query string to select all columns and rows from a table named main_table. The query is then sent to the database using dbSendQuery(), and the fetched data is stored in a data frame using dbFetch(). Clears the result set using dbClearResult() and disconnects from the database using dbDisconnect() to free up resources. The `data_frame` object remains in memory for analysis even after disconnecting from database.

NOTE: Before executing a new query, RUN the dbClearResult(query) instruction to clean memory from the previous query. Additionally, you should run the dbClearResult(query) before closing the database connection. 

For complex queries, we suggest downloading a subset of data to filter and manipulate the data frame obtained using the functions of R. 
```{r}
# Define the SQL query

# Example query 1: this query returns all records from the table main_table
query_string <- "SELECT * FROM main_table"

# Example query 2: this query returns all records where crop is Wheat
# query_string <- "SELECT * FROM main_table WHERE crop = 'Wheat'"

# Example query 3: this query returns all records where crop is Wheat and measurement_method is Remote sensing
# query_string <- "SELECT * FROM main_table WHERE crop = 'Wheat' AND measurement_method = 'Remote sensing'"


# Example query 4: this query returns all records where crop is Wheat and measurement_method is Direct measurement or Proximal sensing
# query_string <- "SELECT * FROM main_table WHERE crop = 'Wheat' 
#                         AND measurement_method = 'Direct measurement' 
#                         OR measurement_method = 'Proximal sensing'"

# Example query 5: this query returns records from columns variable_value and variable_units where crop is Wheat, treatment is Drought and season is Winter 2021-2022
# query_string <- "SELECT variable_value, variable_units FROM main_table WHERE variable_name = 'Leaf SPAD-502 readings' 
#                         AND treatment =  'Drought'
#                         AND season = 'Winter 2021-2022'"

# Send query to the database
query <- dbSendQuery(con, query_string)

# Fetch the data into a data frame
data_frame <- dbFetch(query)

# Clear previous query
dbClearResult(query)

# Disconnect from the database to release resources.
dbDisconnect(con)
```


### Displaying dataframe
Displays the fetched data stored in the `data_frame` object. Data are ready for visualization and analysis!

```{r}
# The results are stored in the data_frame object
head(data_frame)
```